---
output: 
  pdf_document:
    keep_tex: yes
    fig_caption: yes
    number_sections: yes
geometry: margin=2.54cm
title: "Hydrological Characteristics of a Flood Event on the Lower Roanoke River"
subtitle: "https://github.com/YFZeng07/ENV790_WDA_FinalProject.git"
author: "Yingfan Zeng"
fontsize: 12pt
mainfont: Times New Roman

---

\newpage

```{r setup, include=FALSE}
# Set your working directory
getwd()

# Load your packages
library(tidyverse)
library(dplyr)
library(ggpubr)
library(dataRetrieval)
library(zoo)
library(sf)
library(nhdplusTools)
library(viridis)
library(knitr)

# Set your ggplot theme
MyTheme <- theme_bw(base_size = 14) +
  theme(axis.text = element_text(color = "black"),
        legend.position = "top",
        plot.title = element_text(face = "bold", hjust = 0))

# Load your datasets
#Lower Roanoke River Basin map
LRB <- st_read("./Data/Basin_shp/Lower_Roanoke_River_Basin.shp")

#Import the Roanoke River mainstem flowline
#Set Gage1 as the start point
start_point <- st_sfc(st_point(c(Flood_df$dec_long_va[1], 
                                 Flood_df$dec_lat_va[1])), 
                      crs = 4269)
start_comid <- discover_nhdplus_id(start_point)

#Navigate the NLDI network
NLDI <- navigate_nldi(list(featureSource = "comid", featureID = start_comid), 
                      mode = "DM", #downstream mainstem of Gage1
                      distance_km = 1000)

#Extract watershed and flowpath information
subset_file <- tempfile(fileext = ".gpkg")
subset <- subset_nhdplus(comids = as.integer(NLDI$DM$nhdplus_comid),
                         output_file = subset_file,
                         nhdplus_data = "download", 
                         flowline_only = FALSE,
                         return_data = TRUE, overwrite = TRUE)

#Extract the flowline
flowline <- subset$NHDFlowline_Network


```


# Rationale and Research Questions

Flooding is one of the most dangerous natural disasters in the world, affecting societies, economies, and ecosystems. With climate change, floods are increasing in intensity and frequency (Bates, Kundzewicz, Wu, & Palutikof, 2008). Flooding is the main stressor for many ecosystems like forests. Globally speaking, there is a strong link between increased flood frequency and forest loss, as well as the intensity and duration of floods (Bradshaw, Sodhi, PEH, & Brook, 2007). Lower Roanoke River Basin has one of the most biodiverse ecosystems in the Southeast United States. It is a typical southeastern alluvial system, contains the largest natural bottomland hardwood forests in the mid-Atlantic region, and is home to a massive number of fish and wildlife species (U.S. Fish & Wildlife Service, 2014). However, forests on Roanoke floodplains have been threatened by flooding there is a strong negative correlation between the vegetation species composition and tree growth rate (Townsend, 2001). Given the lack of study on hydrology characteristics of the Roanoke River flooding, this project selected a specific flood event to investigate the following questions: 

1. How long was the flood duration along the river?
2. How was the water height change during a flood event? 
3. How did the flood peak move along the river? 


\newpage

# Dataset Information
## Study Area
The Lower Roanoke River Basin is in southern Virginia and northeastern North Carolina in the United States. Located between the John H. Kerr Dam and the Albemarle Sound, this 600000-ha basin contains over 200 km of the Roanoke River, with floodplains of about 5-10 km in width. There is more than 25000 ha of floodplain forests in the basin (Townsend, 2001), which are deeply affected by the floods of the Roanoke River.

## Study Period
The discharge of upstream dams is closely related to downstream flooding in the Lower Roanoke Watershed. There are 3 dams upstream from the study area and they have a great impact on the basin: the John H. Kerr Dam, the Gaston Dam, and the Roanoke Rapids Dam. There is no major dam below the Roanoke Rapids Dam and its maximum discharge capacity is 20,000 cubic feet per second (cfs). Therefore, in this study, a flood event was defined as a continuous period of the discharge of the Roanoke Rapids Dam above 20,000 cfs. Because it had the largest dam discharge of 35000 cfs, I chose the flooding event between February 27 and March 11, 2019 to examine its hydrological characteristics.

## Water Height Data
Water height data from USGS gages along the Lower Roanoke River were obtained to characterize the flooding hydrology for this flood event. There are 8 USGS gages on Roanoke River from Roanoke Rapids to the Albemarle Sound (Table 1, Figure 1). Their daily water height between 2/1/2019 and 4/5/2019 was retrieved to cover the flooding period. 

```{r gage info, include = FALSE}
#Obtain gage information
Gage1_info <- readNWISsite("02080500")[,c(1:3, 7:8)]
Gage2_info <- readNWISsite("0208062765")[,c(1:3, 7:8)]
Gage3_info <- readNWISsite("02081000")[,c(1:3, 7:8)]
Gage4_info <- readNWISsite("02081022")[,c(1:3, 7:8)]
Gage5_info <- readNWISsite("02081028")[,c(1:3, 7:8)]
Gage6_info <- readNWISsite("02081054")[,c(1:3, 7:8)]
Gage7_info <- readNWISsite("02081094")[,c(1:3, 7:8)]
Gage8_info <- readNWISsite("0208114150")[,c(1:3, 7:8)]

#Combine gage information
Info_df <- rbind(Gage1_info, Gage2_info, Gage3_info, Gage4_info,
                 Gage5_info, Gage6_info, Gage7_info, Gage8_info)

```

```{r Table 1, echo = FALSE}

knitr::kable(Info_df, caption = "The 8 USGS gages on the Lower Roanoke River")

```

```{r Figure 1, echo = FALSE, fig.cap = "The Lower Roanoke River, its basin and the 8 USGS gages on it."}
#Map study area 
ggplot() +
  geom_sf(data = LRB) +
  geom_sf(data = Flood_sf, size = 2) +
  geom_sf(data = flowline) +
  MyTheme

```

\newpage

# Exploratory Analysis 
To gain a general understanding of how water height changed during and after the flood event, I plotted the water height for each gage from upstream to downstream. For most of the gages, the data were too volatile for subsequent analysis. Thus, I calculated the 5-day moving average water height to smooth the trend. The comparison between the original data and the 5-day average as well as the basic changes in water level at each gage is shown in Figure 2. 

```{r exploratory analysis, include = FALSE}
#Gage 1: USGS 02080500 ROANOKE RIVER AT ROANOKE RAPIDS, NC===========
#Obtain the data
Gage1 <- readNWISdv(siteNumbers = "02080500",
           parameterCd = "00065", # Water height (feet)
           startDate = "2019-02-01",
           endDate = "2019-04-05")
names(Gage1)[4:5] <- c("Water_Height", "Approval.Code")

#Calculate the 5-day average water height and daily wather height difference
Gage1 <- Gage1 %>%
  mutate(avg_5d = rollmean(Water_Height, k = 5, fill = NA),
         diff = avg_5d - lag(avg_5d))

#Preliminary visualization
Gage1_plot <- ggplot(Gage1, aes(x = Date)) +
  geom_line(aes(y = Water_Height, col = "Original")) +
  geom_line(aes(y = avg_5d, col = "5-day average"), size = 1) +
  labs(title = "Gage 1 - 02080500",
       y = "Water height (ft)") + 
  scale_x_date(date_labels = "%m/%d") +
  scale_color_manual(name = " ", 
                     values = c("Original" = "black", 
                                 "5-day average" = "blue")) +
  MyTheme

#Gage 2: USGS 0208062765 ROANOKE RIVER AT HALIFAX, NC====================
Gage2_info <- readNWISsite("0208062765")[,c(1:3, 7:8)]

#Obtain the data
Gage2 <- readNWISdv(siteNumbers = "0208062765",
                    parameterCd = "00065", # Water height (feet)
                    startDate = "2019-02-01",
                    endDate = "2019-04-05")
names(Gage2)[4:5] <- c("Water_Height", "Approval.Code")

#Calculate the 5-day average water height and daily wather height difference
Gage2 <- Gage2 %>%
  mutate(avg_5d = rollmean(Water_Height, k = 5, fill = NA),
         diff = avg_5d - lag(avg_5d))

#Preliminary visualization
Gage2_plot <- ggplot(Gage2, aes(x = Date)) +
  geom_line(aes(y = Water_Height, col = "Original")) +
  geom_line(aes(y = avg_5d, col = "5-day average"), size = 1) +
  labs(title = "Gage 2 - 0208062765",
       y = "Water height (ft)") + 
  scale_x_date(date_labels = "%m/%d") +
  scale_color_manual(name = " ", 
                     values = c("Original" = "black", 
                                "5-day average" = "blue")) +
  MyTheme

#Gage 3: USGS 02081000 ROANOKE RIVER NEAR SCOTLAND NECK, NC======================
Gage3_info <- readNWISsite("02081000")[,c(1:3, 7:8)]

#Obtain the data
Gage3 <- readNWISdv(siteNumbers = "02081000",
                    parameterCd = "00065", # Water height (feet)
                    startDate = "2019-02-01",
                    endDate = "2019-04-05")
names(Gage3)[4:5] <- c("Water_Height", "Approval.Code")

#Calculate the 5-day average water height and daily wather height difference
Gage3 <- Gage3 %>%
  mutate(avg_5d = rollmean(Water_Height, k = 5, fill = NA),
         diff = avg_5d - lag(avg_5d))

#Preliminary visualization
Gage3_plot <- ggplot(Gage3, aes(x = Date)) +
  geom_line(aes(y = Water_Height, col = "Original")) +
  geom_line(aes(y = avg_5d, col = "5-day average"), size = 1) +
  labs(title = "Gage 3 - 02081000",
       y = "Water height (ft)") + 
  scale_x_date(date_labels = "%m/%d") +
  scale_color_manual(name = " ", 
                     values = c("Original" = "black", 
                                "5-day average" = "blue")) +
  MyTheme

#Gage 4: USGS 02081022 ROANOKE RIVER NEAR OAK CITY, NC===========================
Gage4_info <- readNWISsite("02081022")[,c(1:3, 7:8)]

#Obtain the data
Gage4 <- readNWISdv(siteNumbers = "02081022",
                    parameterCd = "00065", # Water height (feet)
                    startDate = "2019-02-01",
                    endDate = "2019-04-05")
names(Gage4)[4:5] <- c("Water_Height", "Approval.Code")

#Calculate the 5-day average water height and daily wather height difference
Gage4 <- Gage4 %>%
  mutate(avg_5d = rollmean(Water_Height, k = 5, fill = NA),
         diff = avg_5d - lag(avg_5d))

#Preliminary visualization
Gage4_plot <- ggplot(Gage4, aes(x = Date)) +
  geom_line(aes(y = Water_Height, col = "Original")) +
  geom_line(aes(y = avg_5d, col = "5-day average"), size = 1) +
  labs(title = "Gage 4 - 02081022",
       y = "Water height (ft)") + 
  scale_x_date(date_labels = "%m/%d") +
  scale_color_manual(name = " ", 
                     values = c("Original" = "black", 
                                "5-day average" = "blue")) +
  MyTheme

#Gage 5: USGS 02081028 ROANOKE RIVER AT HAMILTON, NC===============================
Gage5_info <- readNWISsite("02081028")[,c(1:3, 7:8)]

#Obtain the data
Gage5 <- readNWISdv(siteNumbers = "02081028",
                    parameterCd = "00065", # Water height (feet)
                    startDate = "2019-02-01",
                    endDate = "2019-04-05")
names(Gage5)[4:5] <- c("Water_Height", "Approval.Code")

#Calculate the 5-day average water height and daily wather height difference
Gage5 <- Gage5 %>%
  mutate(avg_5d = rollmean(Water_Height, k = 5, fill = NA),
         diff = avg_5d - lag(avg_5d))

#Preliminary visualization
Gage5_plot <- ggplot(Gage5, aes(x = Date)) +
  geom_line(aes(y = Water_Height, col = "Original")) +
  geom_line(aes(y = avg_5d, col = "5-day average"), size = 1) +
  labs(title = "Gage 5 - 02081028",
       y = "Water height (ft)") + 
  scale_x_date(date_labels = "%m/%d") +
  scale_color_manual(name = " ", 
                     values = c("Original" = "black", 
                                "5-day average" = "blue")) +
  MyTheme


#Gage 6: USGS 02081054 ROANOKE RIVER AT WILLIAMSTON, NC=====================
Gage6_info <- readNWISsite("02081054")[,c(1:3, 7:8)]

#Obtain the data
Gage6 <- readNWISdv(siteNumbers = "02081054",
                    parameterCd = "00065", # Water height (feet)
                    startDate = "2019-02-01",
                    endDate = "2019-04-05")
names(Gage6)[4:5] <- c("Water_Height", "Approval.Code")

#Calculate the 5-day average water height and daily wather height difference
Gage6 <- Gage6 %>%
  mutate(avg_5d = rollmean(Water_Height, k = 5, fill = NA),
         diff = avg_5d - lag(avg_5d))

#Preliminary visualization
Gage6_plot <- ggplot(Gage6, aes(x = Date)) +
  geom_line(aes(y = Water_Height, col = "Original")) +
  geom_line(aes(y = avg_5d, col = "5-day average"), size = 1) +
  labs(title = "Gage 6 - 02081054",
       y = "Water height (ft)") + 
  scale_x_date(date_labels = "%m/%d") +
  scale_color_manual(name = " ", 
                     values = c("Original" = "black", 
                                "5-day average" = "blue")) +
  MyTheme

#Gage 7: USGS 02081094 ROANOKE RIVER AT JAMESVILLE, NC========================
Gage7_info <- readNWISsite("02081094")[,c(1:3, 7:8)]

#Obtain the data
Gage7 <- readNWISdv(siteNumbers = "02081094",
                    parameterCd = "00065", # Water height (feet)
                    startDate = "2019-02-01",
                    endDate = "2019-04-05")
names(Gage7)[4:5] <- c("Water_Height", "Approval.Code")

#Calculate the 5-day average water height and daily wather height difference
Gage7 <- Gage7 %>%
  mutate(avg_5d = rollmean(Water_Height, k = 5, fill = NA),
         diff = avg_5d - lag(avg_5d))

#Preliminary visualization
Gage7_plot <- ggplot(Gage7, aes(x = Date)) +
  geom_line(aes(y = Water_Height, col = "Original")) +
  geom_line(aes(y = avg_5d, col = "5-day average"), size = 1) +
  labs(title = "Gage 7 - 02081094",
       y = "Water height (ft)") + 
  scale_x_date(date_labels = "%m/%d") +
  scale_color_manual(name = " ", 
                     values = c("Original" = "black", 
                                "5-day average" = "blue")) +
  MyTheme

#Gage 8: USGS 0208114150 ROANOKE RIVER AT NC 45 NR WESTOVER, NC=========================
Gage8_info <- readNWISsite("0208114150")[,c(1:3, 7:8)]

#Obtain the data
Gage8 <- readNWISdv(siteNumbers = "0208114150",
                    parameterCd = "00065", # Water height (feet)
                    startDate = "2019-02-01",
                    endDate = "2019-04-05")
names(Gage8)[4:5] <- c("Water_Height", "Approval.Code")

#Calculate the 5-day average water height and daily wather height difference
Gage8 <- Gage8 %>%
  mutate(avg_5d = rollmean(Water_Height, k = 5, fill = NA),
         diff = avg_5d - lag(avg_5d))

#Preliminary visualization
Gage8_plot <- ggplot(Gage8, aes(x = Date)) +
  geom_line(aes(y = Water_Height, col = "Original")) +
  geom_line(aes(y = avg_5d, col = "5-day average"), size = 1) +
  labs(title = "Gage 8 - 0208114150",
       y = "Water height (ft)") + 
  scale_x_date(date_labels = "%m/%d") +
  scale_color_manual(name = " ", 
                     values = c("Original" = "black", 
                                "5-day average" = "blue")) +
  MyTheme


```

```{r Figure 2, echo = FALSE, fig.cap = "The water height changes during the flood event at each gages"}

#Combine the plots
ggarrange(Gage1_plot, Gage2_plot, Gage3_plot, Gage4_plot, 
          Gage5_plot, Gage6_plot, Gage7_plot, Gage8_plot,
          ncol = 4, nrow = 2)

```


\newpage

# Analysis



## Question 1: <insert specific question here and add additional subsections for additional questions below, if needed>

## Question 2: 

```{r analysis, include = FALSE}

#Gage 1 ==================================
#Find the start date - the first day that the water height is 1 feet higher than the previous day
Gage1_startdate <- Gage1$Date[min(which(diff(Gage1$avg_5d) > 0.5, FALSE))]

#Remove the days before the start date to find the end date
Gage1_subset <- subset(Gage1, Date >= Gage1_startdate)

#Find the end date - the last day of the pulse 
Gage1_enddate <- Gage1_subset$Date[max(which(Gage1_subset$diff < 0))]

#flood duration = end date - start date
Gage1_duration <- as.numeric(Gage1_enddate - Gage1_startdate)

#peak = the date with the max water height 
Gage1_peak <- Gage1_subset$Date[which.max(Gage1_subset$avg_5d)]

#Create a dataframe of just the flood period
Gage1_flood <- Gage1 %>%
  subset(Date >= Gage1_startdate & Date <= Gage1_enddate)

#Calculate the water height increase rate: (max-min)/min
Gage1_flood_rate <- (max(Gage1_flood$avg_5d) - min(Gage1_flood$avg_5d))/min(Gage1_flood$avg_5d)

#Create a list for the flood information
Gage1_summary <- list(Gage1_startdate, Gage1_enddate, Gage1_peak, Gage1_duration, Gage1_flood_rate)

#Gage 2===============================
#Find the start date - the first day that the water height is 1 feet higher than the previous day
Gage2_startdate <- Gage2$Date[min(which(diff(Gage2$avg_5d) > 1, FALSE))]

#Remove the days before the start date to find the end date
Gage2_subset <- subset(Gage2, Date >= Gage2_startdate)

#Find the end date - the last day of the pulse 
Gage2_enddate <- Gage2_subset$Date[max(which(Gage2_subset$diff < 0))]

#flood duration = end date - start date
Gage2_duration <- as.numeric(Gage2_enddate - Gage2_startdate)

#peak = the date with the max water height 
Gage2_peak <- Gage2_subset$Date[which.max(Gage2_subset$avg_5d)]

#Create a dataframe of just the flood period
Gage2_flood <- Gage2 %>%
  subset(Date >= Gage2_startdate & Date <= Gage2_enddate)

#Calculate the water height increase rate: (max-min)/min
Gage2_flood_rate <- (max(Gage2_flood$avg_5d) - min(Gage2_flood$avg_5d))/min(Gage2_flood$avg_5d)

#Create a list for the flood information
Gage2_summary <- list(Gage2_startdate, Gage2_enddate, Gage2_peak, Gage2_duration, Gage2_flood_rate)


#Gage3 ==================================
#Find the start date - the first day that the water height is 1 feet higher than the previous day
Gage3_startdate <- Gage3$Date[min(which(diff(Gage3$avg_5d) > 1, FALSE))]

#Remove the days before the start date to find the end date
Gage3_subset <- subset(Gage3, Date >= Gage3_startdate)

#Find the end date - the last day of the pulse 
Gage3_enddate <- Gage3_subset$Date[max(which(Gage3_subset$diff < 0))]

#flood duration = end date - start date
Gage3_duration <- as.numeric(Gage3_enddate - Gage3_startdate)

#peak = the date with the max water height 
Gage3_peak <- Gage3_subset$Date[which.max(Gage3_subset$avg_5d)]

#Create a dataframe of just the flood period
Gage3_flood <- Gage3 %>%
  subset(Date >= Gage3_startdate & Date <= Gage3_enddate)

#Calculate the water height increase rate: (max-min)/min
Gage3_flood_rate <- (max(Gage3_flood$avg_5d) - min(Gage3_flood$avg_5d))/min(Gage3_flood$avg_5d)

#Create a list for the flood information
Gage3_summary <- list(Gage3_startdate, Gage3_enddate, Gage3_peak, Gage3_duration, Gage3_flood_rate)


#Gage4 ==================================
#Find the start date - the first day that the water height is 1 feet higher than the previous day
Gage4_startdate <- Gage4$Date[min(which(diff(Gage4$avg_5d) > 0.5, FALSE))]

#Remove the days before the start date to find the end date
Gage4_subset <- subset(Gage4, Date >= Gage4_startdate)

#Find the end date - the last day of the pulse 
Gage4_enddate <- Gage4_subset$Date[max(which(Gage4_subset$diff < 0))]

#flood duration = end date - start date
Gage4_duration <- as.numeric(Gage4_enddate - Gage4_startdate)

#peak = the date with the max water height 
Gage4_peak <- Gage4_subset$Date[which.max(Gage4_subset$avg_5d)]

#Create a dataframe of just the flood period
Gage4_flood <- Gage4 %>%
  subset(Date >= Gage4_startdate & Date <= Gage4_enddate)

#Calculate the water height increase rate: (max-min)/min
Gage4_flood_rate <- (max(Gage4_flood$avg_5d) - min(Gage4_flood$avg_5d))/min(Gage4_flood$avg_5d)

#Create a list for the flood information
Gage4_summary <- list(Gage4_startdate, Gage4_enddate, Gage4_peak, Gage4_duration, Gage4_flood_rate)


#Gage5 ==================================
#Find the start date - the first day that the water height is 1 feet higher than the previous day
Gage5_startdate <- Gage5$Date[min(which(diff(Gage5$avg_5d) > 0.3, FALSE))]

#Remove the days before the start date to find the end date
Gage5_subset <- subset(Gage5, Date >= Gage5_startdate)

#Find the end date - the last day of the pulse 
Gage5_enddate <- Gage5_subset$Date[max(which(Gage5_subset$diff < 0))]

#flood duration = end date - start date
Gage5_duration <- as.numeric(Gage5_enddate - Gage5_startdate)

#peak = the date with the max water height 
Gage5_peak <- Gage5_subset$Date[which.max(Gage5_subset$avg_5d)]

#Create a dataframe of just the flood period
Gage5_flood <- Gage5 %>%
  subset(Date >= Gage5_startdate & Date <= Gage5_enddate)

#Calculate the water height increase rate: (max-min)/min
Gage5_flood_rate <- (max(Gage5_flood$avg_5d) - min(Gage5_flood$avg_5d))/min(Gage5_flood$avg_5d)

#Create a list for the flood information
Gage5_summary <- list(Gage5_startdate, Gage5_enddate, Gage5_peak, Gage5_duration, Gage5_flood_rate)


#Gage6 ==================================
#Find the start date - the first day that the water height is 1 feet higher than the previous day
Gage6_startdate <- Gage6$Date[min(which(diff(Gage6$avg_5d) > 0.1, FALSE))]

#Remove the days before the start date to find the end date
Gage6_subset <- subset(Gage6, Date >= Gage6_startdate)

#Find the end date - the last day of the pulse 
Gage6_enddate <- Gage6_subset$Date[max(which(Gage6_subset$diff < 0))]

#flood duration = end date - start date
Gage6_duration <- as.numeric(Gage6_enddate - Gage6_startdate)

#peak = the date with the max water height 
Gage6_peak <- Gage6_subset$Date[which.max(Gage6_subset$avg_5d)]

#Create a dataframe of just the flood period
Gage6_flood <- Gage6 %>%
  subset(Date >= Gage6_startdate & Date <= Gage6_enddate)

#Calculate the water height increase rate: (max-min)/min
Gage6_flood_rate <- (max(Gage6_flood$avg_5d) - min(Gage6_flood$avg_5d))/min(Gage6_flood$avg_5d)

#Create a list for the flood information
Gage6_summary <- list(Gage6_startdate, Gage6_enddate, Gage6_peak, Gage6_duration, Gage6_flood_rate)


#Gage7 ==================================
#Find the start date - the first day that the water height is 1 feet higher than the previous day
Gage7_startdate <- Gage7$Date[min(which(diff(Gage7$avg_5d) > 0.06, FALSE))]

#Remove the days before the start date to find the end date
Gage7_subset <- subset(Gage7, Date >= Gage7_startdate)

#Find the end date - the last day of the pulse 
Gage7_enddate <- Gage7_subset$Date[max(which(Gage7_subset$diff < 0))]

#flood duration = end date - start date
Gage7_duration <- as.numeric(Gage7_enddate - Gage7_startdate)

#peak = the date with the max water height 
Gage7_peak <- Gage7_subset$Date[which.max(Gage7_subset$avg_5d)]

#Create a dataframe of just the flood period
Gage7_flood <- Gage7 %>%
  subset(Date >= Gage7_startdate & Date <= Gage7_enddate)

#Calculate the water height increase rate: (max-min)/min
Gage7_flood_rate <- (max(Gage7_flood$avg_5d) - min(Gage7_flood$avg_5d))/min(Gage7_flood$avg_5d)

#Create a list for the flood information
Gage7_summary <- list(Gage7_startdate, Gage7_enddate, Gage7_peak, Gage7_duration, Gage7_flood_rate)


#Gage8 ==================================
#Gage8 data is funky so I refined the time frame 
#Obtain the data
Gage8_update <- readNWISdv(siteNumbers = "0208114150",
                    parameterCd = "00065", # Water height (feet)
                    startDate = "2019-02-20",
                    endDate = "2019-04-15")
names(Gage8_update)[4:5] <- c("Water_Height", "Approval.Code")

#Calculate the 5-day average water height and daily wather height difference
Gage8_update <- Gage8_update %>%
  mutate(avg_5d = rollmean(Water_Height, k = 5, fill = NA),
         diff = avg_5d - lag(avg_5d))

ggplot(Gage8_update, aes(x = Date, y = avg_5d)) +
  geom_line()

#Find the start date - the first day that the water height is 1 feet higher than the previous day
Gage8_startdate <- Gage8_update$Date[min(which(diff(Gage8_update$avg_5d) > 0.05, FALSE))]

#Remove the days before the start date to find the end date
Gage8_subset <- subset(Gage8_update, Date >= Gage8_startdate)

#Find the end date - the last day of the pulse 
Gage8_enddate <- Gage8_subset$Date[max(which(Gage8_subset$diff < 0))]

#flood duration = end date - start date
Gage8_duration <- as.numeric(Gage8_enddate - Gage8_startdate)

#peak = the date with the max water height 
Gage8_peak <- Gage8_subset$Date[which.max(Gage8_subset$avg_5d)]

#Create a dataframe of just the flood period
Gage8_flood <- Gage8_update %>%
  subset(Date >= Gage8_startdate & Date <= Gage8_enddate)

#Calculate the water height increase rate: (max-min)/min
Gage8_flood_rate <- (max(Gage8_flood$avg_5d) - min(Gage8_flood$avg_5d))/min(Gage8_flood$avg_5d)

#Create a list for the flood information
Gage8_summary <- list(Gage8_startdate, Gage8_enddate, Gage8_peak, Gage8_duration, Gage8_flood_rate)



#Create a data freame to gather all information for all 8 gages================

#Create an empty data frame for flood information
Summary_df <- data.frame(Start = as.Date(as.character()),    
                    End = as.Date(as.character()),
                    Peak = as.Date(as.character()),
                    Duration = numeric(),
                    IncreasedPercent = numeric(),
                    stringsAsFactors = FALSE)

#Add flood information
Summary_df[1,] <- Gage1_summary
Summary_df[2,] <- Gage2_summary
Summary_df[3,] <- Gage3_summary
Summary_df[4,] <- Gage4_summary
Summary_df[5,] <- Gage5_summary
Summary_df[6,] <- Gage6_summary
Summary_df[7,] <- Gage7_summary
Summary_df[8,] <- Gage8_summary

#Combine gage information
Info_df <- rbind(Gage1_info, Gage2_info, Gage3_info, Gage4_info,
                 Gage5_info, Gage6_info, Gage7_info, Gage8_info)

#Combine gage information and flood information
Flood_df <- cbind(Info_df, Summary_df)


```


```{r Table 2, echo = FALSE}

knitr::kable(Info_df, caption = "The 8 USGS gages on the Lower Roanoke River")

```

\newpage

# Summary and Conclusions


\newpage

# References
Bates, B. C., Kundzewicz, Z. W., Wu, S., & Palutikof, J. P. (2008). Climate Change and Water. Technical Paper of the Intergovernmental Panel on Climate Change, IPCC Secretariat: Geneva, Switzerland. The American Midland Naturalist, 168(1).

Bradshaw, C. J., Sodhi, N. S., PEH, K. S. H., & Brook, B. W. (2007). Global evidence that deforestation amplifies flood risk and severity in the developing world. Global Change Biology, 13(11), 2379-2395.

U.S. Fish & Wildlife Service. (2014). National Wildlife Refuge: Roanoke River. https://www.fws.gov/refuge/Roanoke_River/wildlife_and_habitat/index.html

Townsend, P. A. (2001). Relationships between vegetation patterns and hydroperiod on the Roanoke River floodplain, North Carolina. Plant Ecology, 156(1), 43-58.

